### 2025-07-08 factor-analysis.qmd
20250708 - tweaks

@@

psych::fa.parallel(
  x = dataMerged[faVars],
  fm = "minres", # errors out when using "ml"
  fa = "fa"
)
```

@@

screeDataEFA <- nFactors::nScree(
  x = cor(
    dataMerged[faVars],
    use = "pairwise.complete.obs"),
  model = "factors")

@@

psych::vss(
  dataMerged[faVars],
  rotate = "oblimin",
  fm = "minres") # errors out when using "mle"
```

@@

psych::nfactors(
  dataMerged[faVars],
  rotate = "oblimin",
  fm = "minres") # errors out when using "ml"
```

@@

efaOblique_fit <- lavaan::efa(
  data = dataMerged,
  ov.names = faVars,
  nfactors = 1:9,
  rotation = "geomin",
  missing = "ML",
  estimator = "MLR",
  meanstructure = TRUE
)
```

@@

cfaModel1_fit <- lavaan::cfa(
  cfaModel1_syntax,
  data = dataMerged,
  missing = "ML",
  estimator = "MLR",
  std.lv = TRUE)
```

>>>>

#### Standardize Variables {#sec-pcaPrepareDataStandardize}

Although standardizing variables is not strictly necessary in factor analysis, it can be helpful to prevent some variables from having considerably larger variances than others.

```{r}
dataForFA <- dataMerged
dataForFA[faVars] <- scale(dataForFA[faVars])
```

@@

psych::fa.parallel(
  x = dataForFA[faVars],
  fm = "minres", # errors out when using "ml"
  fa = "fa"
)
```

@@

screeDataEFA <- nFactors::nScree(
  x = cor(
    dataForFA[faVars],
    use = "pairwise.complete.obs"),
  model = "factors")

@@

psych::vss(
  dataForFA[faVars],
  rotate = "oblimin",
  fm = "minres") # errors out when using "mle"
```

@@

psych::nfactors(
  dataForFA[faVars],
  rotate = "oblimin",
  fm = "minres") # errors out when using "ml"
```

@@

efaOblique_fit <- lavaan::efa(
  data = dataForFA,
  ov.names = faVars,
  nfactors = 1:9,
  rotation = "geomin",
  missing = "ML",
  estimator = "MLR",
  cluster = "player_id",
  meanstructure = TRUE
)
```

@@

cfaModel1_fit <- lavaan::cfa(
  cfaModel1_syntax,
  data = dataForFA,
  missing = "ML",
  estimator = "MLR",
  std.lv = TRUE,
  cluster = "player_id",
  em.iter.max = 2000000)
```

### 2025-07-08 machine-learning.qmd
20250708 - tweaks

The most accurate of models all explained less than half of the variance in players' future fantasy points.
Thus, there remains considerable variance to be explained.
The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.
It is also consistent with the conclusion from many other domains in psychology that human behavior is challenging to predict [@PetersenPrinciplesPsychAssessment].

>>>>

The most accurate of models all explained less than half of the variance in players' future fantasy points.
Thus, there remains considerable variance to be explained.
The considerable unexplained variance suggests that we are missing key variables that are important predictors of future fantasy performance.
It is also consistent with the conclusion from many other domains in psychology that human behavior is challenging to predict [@PetersenPrinciplesPsychAssessment].
We could likely improve our predictive accuracy by including projections, but it is unclear what information projections are or are not based on.
For instance, if projections already account for the prior season's statistics and fantasy points, then including both may not buy you much in terms of increased predictive accuracy.
